DRAFT_Cleaning

### 2025-09-03 CSAT_Rates.Rmd
catching up for capstone

```{r}
# Set folder path
folder_path <- "C:/workspace/CSAT/CSAT_distribution/output/CEMVR/19990101__to__20241231"

@@

@@

combined_data_clean<-combined_data|>
  mutate(Before_date = ymd(SurveyDateBefore),
         After_date = ymd(SurveyDateAfter))|>
  mutate(DateDiff = as.numeric(After_date - Before_date))|>
  mutate(DailyShoalingRate_ftperday = AnnualShoalingRate_ftperyr/DateDiff,
         DailyShoalingVolume_Cyperday = AnnualShoalingVolume_CYperyr/DateDiff,
         VolChange_Cyperday = NetVolumeChange_CY/DateDiff)
#  filter(DailyShoalingVolume_Cyperday < 12500000 & DailyShoalingRate_ftperday < 900)
```

@@

```
Let's only look at rate differences calculated within a year 
```{r}
combined_data_year <- combined_data_clean|>
  mutate(yr_diff = abs(year(After_date) - year(Before_date)))|>
  filter(yr_diff <1)|>
  filter(pool != "AL")|>
  filter(pool != "CS")
```

```{r}
ggplot(combined_data_year, aes(x = After_date, y = AnnualShoalingVolume_CYperyr, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(NA,2000000)+
  facet_wrap(~ river)

ggplot(combined_data_year, aes(x = After_date, y = AnnualShoalingRate_ftperyr, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(NA, 275)+
  facet_wrap(~ river)

ggplot(combined_data_year, aes(x = After_date, y = NetVolumeChange_CY, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(NA, 275)+
  facet_wrap(~ river)

ggplot(combined_data_year, aes(x = Before_date, y = NetVolumeChange_CY, color = river)) +
  geom_point() +
  theme(legend.position = "bottom")+
  geom_line(alpha = 0.5) +
  labs(title = "Net Volume Change in Sediment (CY) Over Time By Pool",
       x = "Survey Date",
       y = "Net Volume Change (CY)",
       color = "River") +
  theme_minimal() +
    facet_wrap(~ pool, scales = "free_y")
```

>>>>

```{r}
# Set folder path
folder_path <- "C:/workspace/CSAT/CSAT_distribution/output/CEMVR/19990101_to_20241231_Aug"

@@

```{r}
surveyUpdate_table<-read_delim("SurveyUpdate.txt")
dredge_data <-  read.csv("Dredge_Event_data.csv")

dredge_data<- dredge_data |>
  filter(!is.na(POSTHYDROSURVEY_ID))|>
  filter(POSTHYDROSURVEY_ID != "")

dredge_survey_joined <- left_join(surveyUpdate_table, dredge_data, 
                                  by = join_by(SurveyID == POSTHYDROSURVEY_ID), 
                                  relationship = "many-to-many", keep = TRUE)|>
  mutate(SurveyType, case_when(
    !is.na(POSTHYDROSURVEY_ID) ~ "AD",
    TRUE ~ SurveyType))

@@

bad_data <- combined_data|>
  filter(abs(AnnualShoalingRate_ftperyr) >= 30)

combined_data_clean<-combined_data|>
  mutate(Before_date = ymd(SurveyDateBefore),
         After_date = ymd(SurveyDateAfter))|>
  mutate(DateDiff = as.numeric(After_date - Before_date))|>
  mutate(DailyShoalingRate_ftperday = AnnualShoalingRate_ftperyr/DateDiff,
         DailyShoalingVolume_Cyperday = AnnualShoalingVolume_CYperyr/DateDiff,
         VolChange_Cyperday = NetVolumeChange_CY/DateDiff)
```

@@

Let's only look at rate differences calculated within a year 
```{r}
combined_data_year <- combined_data_clean|>
  mutate(yr_diff = abs(year(After_date) - year(Before_date)))|>
  filter(yr_diff <1)|>
  filter(pool != "AL")|>
  filter(pool != "CS")
```

```{r}
outliers_zscore <- combined_data_clean |>
  group_by(pool) |>
  mutate(score = as.numeric(scale(AnnualShoalingRate_ftperyr))) |>
  filter(abs(score) > 3) |>  # Points more than 3 standard deviations from mean
  arrange(pool, desc(abs(score)))|>
  filter(abs(AnnualShoalingRate_ftperyr) >= 30)

write_csv(outliers_zscore,"Outliers_shoalingRateOmitted.csv")
```

```{r}

ggplot(combined_data_clean, aes(x = After_date, y = AnnualShoalingVolume_CYperyr, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(NA,2000000)+
  facet_wrap(~ river)

ggplot(combined_data_clean, aes(x = After_date, y = AnnualShoalingRate_ftperyr, colour = river)) +
  geom_point() + 
  geom_line()+
   labs(title = "Annual Shoaling Rate (ft/yr) By Pool",
       x = "Survey Date",
       y = "Annual Shoaling Rate (ft/yr)",
       color = "River") +
 # ylim(-30, 30)+
  facet_wrap(~ pool)

ggplot(combined_data_clean, aes(x = Before_date, y = NetVolumeChange_CY, color = river)) +
  geom_point() +
  theme(legend.position = "bottom")+
  geom_line(alpha = 0.5) +
  labs(title = "Net Volume Change in Sediment (CY) Over Time By Pool",
       x = "Survey Date",
       y = "Net Volume Change (CY)",
       color = "River") +
  theme_minimal() +
    facet_wrap(~ pool, scales = "free_y")
```

### 2025-09-03 Draft_Cleaning.Rmd
catching up for capstone

###Cross Validation of Model
```{r}
library(progress)
pb <- progress_bar$new(total = n_outer_folds)
set.seed(123)
n = nrow(gage_xgb_scaled)
n_outer_folds = 5

@@

gage_xgb_scaled_clean <- gage_xgb_scaled|>
  select(-lead_7,-lead_14)

@@

# Optionally remove future-looking targets
gage_scaled_clean <- gage_scaled_clean |> select(-lead_7, -lead_14)

@@

# Optional: remove future-looking columns
gage_scaled_clean <- gage_scaled_clean |> select(-lead_7, -lead_14)

>>>>

###Cross Validation of Model
```{r}
library(progress)
n_outer_folds = 5
pb <- progress_bar$new(total = n_outer_folds)
set.seed(123)
n = nrow(gage_xgb_scaled)

@@

gage_xgb_scaled_clean <- gage_xgb_scaled

@@

# Optionally remove future-looking targets
gage_scaled_clean <- gage_scaled_clean

@@

# Optional: remove future-looking columns
gage_scaled_clean <- gage_scaled_clean
